\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-02-05 doc/paper.tex
updating from doc_v2

  \author{
        Ganesh Krishnan \thanks{The authors gratefully acknowledge \ldots{}} \\
    Department of Statistics, Iowa State University\\
     and \\     Heike Hofmann \\
    Department of Statistics and csafe, Iowa State University\\
      }
  \maketitle
} \fi

@@

\section{Introduction}\label{introduction}

\subsection{Problem statement and
motivation}\label{problem-statement-and-motivation}

\hh{'considered to be unique' by AFTE - that is right, but we should not use this argument, and I don't think we have to. It's better to focus on the problem at hand: striation marks are used to determine same source for bullets. XXX we could include some figures here XXX Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner. One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}. There are different approaches used to achieve more objectivity and assess error rates. XXX here we need some examples of approaches XXX
For toolmarks, the Chumbley score has been suggested to determine same-source. XXX summary of results from original paper and the adjustment XXX
In this paper we are interested in how this method has to be adjusted to best be applied to the smaller marks on bullets. }
\gk{done}

Striation marks are used to determine same source for bullets. Current
standards ask for a visual inspection of the marks under a comparison
microscope by a firearms examiner. One of the problems raised by the NAS
report in 2009 is the subjectivity involved in this comparison and the
need for determining error rates \citep{NAS:2009}. The issue of
subjectivity of a firearms examiner has been reviewed by many authors
where things like scientific principle testability and error rates are
considered to be defining aspects for an objective analysis such that
without identifying the uncertainities associated with a method of
comparison, it is inconclusive to regards an analysis complete. This
means that with the usual method of visual comparison that is adopted by
firearms examiners, lies the problem of lack of any systemized
quantification and as such there seems no certain way of associating it
with any kind of mathematical probability. Since determining error rates
has been duly noted as a fundamental problem in forensic science
\citep{NAS:2009}, there is a need of methods that address this problem.
Some of the methods that have been used to estimate error rates in
firearm examinations include Automatic cartridge case comparison where
same source identification of cartridge cases and associated error rates
have been provided by \citep{riva}, although as noted by \citep{aoas},
in this case alignments of striae involves roation of planes, which
cannot be generalized for bullets. In case of methods that use machine
learning algorithms some methods which are bootsrap based methods often
tend to give over-estimated error rates \citep{efron}, but there are
alternate error estimation techniques as described by \citep{aoas},
\citep{efron} and \citep{vorburger2016} which give better results.

@@

\citep{chumbley} found that the algorithm gives false-positives in
slightly over 2 cases (or 2\%) of the time and about 9 false-negatives
(or 9\%) for every 100 comparisons. \citep{hadler} on the other hand, in
the improved version of the algorithm first described by
\citep{chumbley} found false-positives for 0 cases and 3 false-negatives
(or 6\%) for 50 comparisons of knwon matches.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6.png}

\begin{center}\includegraphics[width=\textwidth]{figures/unnamed-chunk-1-1} \end{center}

\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification.}

@@

\subsection{Existing solution for toolmarks: Chumbley
Score}\label{existing-solution-for-toolmarks-chumbley-score}

\hh{XXX anything to do with motivation has to come first - move this to the first paragraph of the intro XXX}
\hh{XXX data and results come after the intro to chumbley.}
\hh{XXX start with defining the data, i.e. we are working with random processes $x(t)$ and $y(t)$ ... you should also go into the coarseness parameter here - that is a pre-processing step before the optimization. }
\gk{done}

The Chumbley score algorithm takes input as two vectorized processes
\(x(t_1)\) and \(y(t_2)\) which denote two sets of marks or striae.
These marks or striae are potentially from two different bullets or two
different toolmarks whose source needs to be identified as being same or
different. The marks or striae are indexed by the pixel location where
\(t_1\) is for the first striae referred to as \(x\) and \(t_2\) is for
the second striae which is referred as \(y\). \(t_1\) and \(t_2\) need
not be the same but are usually of similar lenghtsThe similarity is then
judged by the algorithm on the basis of cross-correlation of a fixed and
constant number consecutive pixels say \(k\) taken from the two indexed
marks \(x(t_1)\) and \(y(t_2)\) such that in theory \(k\) remains
smaller that length of the two striae or marks which is the same as
\(t_1\) and \(t_2\). Depending on the what stage of the algorithm we are
in, matching of different pixel lengths and locations is done, which at
the end effectively compares all possible windows that would guarantee
in quantifying the two marks or striae as coming from the same source or
not, which also lets us assess the error rates by checking for a large
number of cases.

@@

A maximum correlation statistic is used to identify the region of best
agreement, with the maximum usually seen being near 1 for both cases
which is what we intuitively expect for matches, and something which we
do not intuitively expect for non matches
\gk{done}\hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}

\hh{XXX the process you describe below is the cross-correlation of two series $x(t)$ and $y(t)$.}

First, there are a very large number of cross-correlations calculated
for the two series of striae or marks \(x(t_1)\) and \(y(t_2)\), for eg
if the window of optimization was defined as 200 and toolmark pixel
length is around 1000 then we have 1200-(200-1) = 1001. Can be thought
of ordered arrangements without replacement where n = 1001 and r = 1.

@@

\begin{verbatim}
  Classification Match Non-Match
1          Match    47         3
2      Non-Match     0        50
\end{verbatim}

@@

Let ns and nd be the number of same shift and different shift windows
\(N = n_{s} + n_{d}\)

The mann whitney U statistic is given by
\(U =\sum^{ns}_{i=1}R_{s}\left( i\right)\)

@@

\(\overline{U}= \dfrac{U-M}{\sqrt{V}}\)

@@

\(M = n_{s}\left(\dfrac{N+1}{2}\right)\)

@@

\(V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}\)

\subsection{\texorpdfstring{Potential limitations in the application to
striation marks on bullet lands: smaller in width and curved - need to
adjust parameter settings; something similar is done in
\citet{afte-chumbley} for toolmark comparisons of slip-joint
pliers.}{Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in  for toolmark comparisons of slip-joint pliers.}}\label{potential-limitations-in-the-application-to-striation-marks-on-bullet-lands-smaller-in-width-and-curved---need-to-adjust-parameter-settings-something-similar-is-done-in-for-toolmark-comparisons-of-slip-joint-pliers.}

Also when in the optimization step, the Window of optimization for
bullets will be shorter as the signatures are smaller. The idea is to
keep the number of windows of optimization sufficiently large, which
means shorter Trace segments (partition of signature or toolmark with
length = size of window of optimization) that lets us compare smaller
segments of one signature to another.

This introduces a problem as, if we go too small in the window of
optimization, the unique features of the trace segments are lost and
seem similar, while too large sizes vastly reduces the weight of small
features that would otherwise uniquely classify a signature and hence
identify the region of agreement.

Thus the Window of Optimization has a direct influence on whether we are
Falsely rejecting the null when it is true (Type I) or Falsely accepting
the null when it is false (Type II) making identification of the optimum
size of window of optimization very important.

\subsection{Using cross-validation setup to identify appropriate
parameter settings for (a) signatures and (b) profiles
directly}\label{using-cross-validation-setup-to-identify-appropriate-parameter-settings-for-a-signatures-and-b-profiles-directly}

Following on similar lines to the setup of toolamrks, the first step
here is to first identify what difference does different window sizes of
optimization and the validation step have, when adapting the toolmark
method to bullets.

The marking made on bullets are smaller than toolmarks and is also less
wider. The idea is to find out possible areas of error while adapting
the score based method proposed for toolmarks.

Bullet signatures being compared at this time are from the Hamby 44 and
Hamby 252 data, which have the best set of known-matched and known
non-matches.

Bullet signatures are chosen by 1. Filtering out Land\_id for Profiles
from the Hamby 44 and Hamby 252 data and removing all NA values 2.
run\_id = 3 is chosen, Seems like the signatures generated from this
run\_id give the closest match. Different run\_id's have some different
settings for generating the signatures (The level of smoothing does not
seem to be one of them)

The bullet signatures when generated already included the LOWESS
smoothing. Therefore, the coarseness factor is set to 1 while running
the chumbley\_non\_random() which generates the same\_shift,
different\_shift, U-Stat and P\_value parameters.

window Optimization 320 Window Validation 50

\begin{verbatim}
       match
signif  FALSE  TRUE
  FALSE 77564   412
  TRUE   4022   674
\end{verbatim}

window Optimization 120 Window Validation 50

\begin{verbatim}
       match
signif  FALSE  TRUE
  FALSE 79249   386
  TRUE   4523   854
\end{verbatim}

window Optimization 80 Window Validation 50

\begin{verbatim}
       match
signif  FALSE  TRUE
  FALSE 78817   450
  TRUE   4680   775
\end{verbatim}

@@

\subsection{Signatures}\label{signatures}

We used the adjusted Chumbley method as proposed in \citet{hadler} and
implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
pairwise land-to-land comparisons of the Hamby scans (a total of 85,491
comparisons) in the following manner:

\caption{\label{tab:unnamed-chunk-5}Overview of parameter settings used for optimization and validation windows for bullet land signatures.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr}
\toprule
wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 50\\
\bottomrule
\end{tabular}}
\end{table}

\subsubsection{Signatures}\label{signatures-1}

Figure \ref{fig:type2} gives an overview of type II error rates observed
when varying the window size in the optimization step.

@@

Figure \ref{fig:type1} compares nominal (fixed) type I error and
actually observed type I errors for the parameter settings in table XXX.
With an increasing size of the window used in the optimization step the
observed type I error rate decreases (slighty). This might be related to
the increasing number of tests that fail for larger window sizes, in
particular for non-matching striae (see fig \ref{fig:missings}).

@@

\caption{Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$. The horizontal line in each facet indicates the nominal type I error rate.  As the optimization window increase the observed type I error rate gets smaller. A smaller validation window tends to be associated with a higher type I error rate.}\label{fig:type1}
\end{figure}

@@

\caption{Number of failed tests by the window optimization size, wo, and ground truth. Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source.}\label{fig:missings}
\end{figure}

Figure \ref{fig:missings} gives an overview of the number of failed
tests, i.e.~tests in which a particular parameter setting did not return
a valid result. This happens, when the shift to align two signatures is
so large, that the remaining overlap is too small to accommodate windows
for validation. The problem is therefore exacerbated by a larger
validation window. Figure \ref{fig:missings} also shows that the number
of failed tests is approximately linear in the size of the optimization
window.

@@

\begin{center}\includegraphics[width=\textwidth]{figures/wo120-1} \end{center}

Coming back to the database of Hamby-44 and Hamby-252 datasets, the
run\_id = 3 was used when applying the chumbley algorithm on the
signatures. The run\_id not only defines the level of smoothing but also
signifies the chosen height at which the profiles were selected
initially. Another important aspect is the range of horizontal values
(which is referred to as the y values in \citet{aoas}) in the
signatures. Thes have already been pre-processed in the database to not
include any grooves.

The idea therefore is to first use these raw values of the profile
directly in the chumbley algorithm, and see how the algorithm performs
for different coarseness values (smoothing parameter as referred in the
function LOWESS used in the chumbley algorithm)

>>>>

  \author{
        Ganesh Krishnan \thanks{The authors gratefully acknowledge \ldots{}} \\
    Department of Statistics, Iowa State University\\
     and \\     Heike Hofmann \\
    Department of Statistics and CSAFE, Iowa State University\\
      }
  \maketitle
} \fi

@@

\tableofcontents
\newpage

\section{Introduction and Background}\label{introduction-and-background}

\subsection{Motivation}\label{motivation}

Striation marks are used to determine same source for bullets. Current
standards ask for a visual inspection of the marks under a comparison
microscope by a firearms examiner. One of the problems raised by the NAS
report in 2009 is the subjectivity involved in this comparison and the
need for determining error rates \citet{NAS:2009}. The issue of
subjectivity of a firearms examiner has been reviewed by many authors
where things like scientific principle testability and error rates are
considered to be defining aspects for an objective analysis such that
without identifying the uncertainities associated with a method of
comparison, it is inconclusive to regards an analysis complete. This
means that with the usual method of visual comparison that is adopted by
firearms examiners, lies the problem of lack of any systemized
quantification and as such there seems no certain way of associating it
with any kind of mathematical probability. Since determining error rates
has been duly noted as a fundamental problem in forensic science
\citet{NAS:2009}, there is a need of methods that address this problem.
Some of the methods that have been used to estimate error rates in
firearm examinations include Automatic cartridge case comparison where
same source identification of cartridge cases and associated error rates
have been provided by \citet{riva}, although as noted by \citet{aoas},
in this case alignments of striae involves roation of planes, which
cannot be generalized for bullets. In case of methods that use machine
learning algorithms some methods which are bootsrap based methods often
tend to give over-estimated error rates \citet{efron}, but there are
alternate error estimation techniques as described by \citet{aoas},
\citet{efron} and \citet{vorburger2016} which give better results.

@@

\citet{chumbley} found that the algorithm gives false-positives in
slightly over 2 cases (or 2\%) of the time and about 9 false-negatives
(or 9\%) for every 100 comparisons. \citet{hadler} on the other hand, in
the improved version of the algorithm first described by
\citet{chumbley} found false-positives for 0 cases and 3 false-negatives
(or 6\%) for 50 comparisons of knwon matches.

\begin{table}[!h]
\centering
\begin{tabular}{lrr}
\toprule
Classification & Match & Non-Match\\
\midrule
Match & 47 & 3\\
Non-Match & 0 & 50\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Potential limitations of the Chumbley Score adaptation to
bullets}\label{potential-limitations-of-the-chumbley-score-adaptation-to-bullets}

Also when in the optimization step, the Window of optimization for
bullets will be shorter as the signatures are smaller. The idea is to
keep the number of windows of optimization sufficiently large, which
means shorter Trace segments (partition of signature or toolmark with
length = size of window of optimization) that lets us compare smaller
segments of one signature to another. This introduces a problem as, if
we go too small in the window of optimization, the unique features of
the trace segments are lost and seem similar, while too large sizes
vastly reduces the weight of small features that would otherwise
uniquely classify a signature and hence identify the region of
agreement.

Thus the Window of Optimization has a direct influence on whether we are
Falsely rejecting the null when it is true (Type I) or Falsely accepting
the null when it is false (Type II) making identification of the optimum
size of window of optimization very important. This raises important
questions about what parameter settings need to be chosen for bullet
land comparison. Something similar is done in \citet{afte-chumbley} for
toolmark comparisons of slip-joint pliers where optimum window sizes are
determined. There is a need to figure out the best parameter settings
which minimizes the errors for unsmoothed markings and pre-processed
signatures as determined by \citet{aoas}. An analysis of these error
rates and comparison with other methods will help us understand the
adaptibility of the chumbley score to bullets.

\hh{figure \ref{fig:rgl} needs to be mentioned somewhere in the write-up}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}

\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom). The dotted lines connect some peaks visible in both visualizations.}

@@

\section{The Chumbley Score Test}\label{the-chumbley-score-test}

The Chumbley score algorithm takes input as two vectorized processes
\(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
which denote two sets of marks or striae. Here the processes are of the
form \(z(s)\) which is a spatial process for some \(t\), while
\(z(s_1)\) is the realization of the process in \(s_1\). The \(t_1\) and
\(t_2\) can be better understood as equally spaced pixel locations for
the two marks under consideration, where a `pixel' is the resolution of
the confocal light microscope. In the case for bullet signatures a pixel
corresponds to 0.645 microns. These marks or striae are potentially from
two different bullets or two different toolmarks whose source needs to
be identified as being same or different. The marks or striae are
indexed by the pixel location where \(t_1\) is for the first striae
referred to as \(x\) and \(t_2\) is for the second striae which is
referred as \(y\). \(T_1\) and \(T_2\) being the respective lengths of
the markings, need not be the same but are usually of similar lenghtsThe
similarity is then judged by the algorithm on the basis of
cross-correlation of a fixed and constant number consecutive pixels say
\(k\) taken from the two indexed marks \(x(t_1)\) and \(y(t_2)\) such
that in theory \(k\) remains smaller that length of the two striae or
marks which is the same as \(t_1\) and \(t_2\). Depending on the what
stage of the algorithm we are in, matching of different pixel lengths
and locations is done, which at the end effectively compares all
possible windows that would guarantee in quantifying the two marks or
striae as coming from the same source or not, which also lets us assess
the error rates by checking for a large number of cases.

@@

A maximum correlation statistic is used to identify the region of best
agreement, with the maximum usually seen being near 1 for both cases
which is what we intuitively expect for matches, and something which we
do not intuitively expect for non matches

First, there are a very large number of cross-correlations calculated
for the two series of striae or marks \(x(t_1)\) and \(y(t_2)\), for eg
if the window of optimization was defined as 200 and toolmark pixel
length is around 1000 then we have 1200-(200-1) = 1001.

@@

@@

Let ns and nd be the number of same shift and different shift windows
\[N = n_{s} + n_{d}\]

The mann whitney U statistic is given by
\[U =\sum^{ns}_{i=1}R_{s}\left( i\right)\]

@@

\[\overline{U}= \dfrac{U-M}{\sqrt{V}}\]

@@

\[M = n_{s}\left(\dfrac{N+1}{2}\right)\]

@@

\[V = \dfrac{n_{s}n_{d}}{N\left(N-1\right)}\left[\Sigma^{n_{s}}R_{s}\left(i\right)^{2}+\Sigma^{n_{d}}R_{d}\left(j\right)^{2}\right] -\dfrac{n_{s} n_{d}\left(N+1\right)^{2}}{4\left(N-1\right)}\]

\subsection{Testing setup}\label{testing-setup}

Following on similar lines to the setup of toolmarks, the first step
here is to first identify what difference does different window sizes of
optimization and the validation step have, when adapting the toolmark
method to bullets.

The marking made on bullets are smaller than toolmarks and is also less
wider. The idea is to find out possible areas of error while adapting
the score based method proposed for toolmarks, using cross-validation
setup to identify appropriate parameter settings for (a) signatures and
(b) profiles directly

\subsubsection{Signatures}\label{signatures}

Bullet signatures being compared at this time are therefore from the
Hamby 44 and Hamby 252 data. The database setup and pre-processing
system used for choosing the Bullet signatures are as described by
\citet{aoas}. In order to choose the bullet signatures we first filter
out Land\_id for Profiles from the Hamby 44 and Hamby 252 data and
remove all NA values. Then run\_id = 3 is chosen as the signatures
generated from this run\_id give the closest match. Different run\_id's
have some different settings for generating the signatures.(The level of
smoothing does not seem to be one of them)

The bullet signatures when generated by this process already includes a
loess smoothing. Therefore, the coarseness factor is set to 1 while
running the chumbley non random algorithm for comparing different
optimization windows.The algorithm generates the same\_shift,
different\_shift, U-Stat and P\_value parameters which are then used to
calculate the errors associated with different sets of window sizes.

Similar to signatures the run\_id = 3 was used when applying the
chumbley algorithm using the database setup given by \citet{aoas} of
Hamby-44 and Hamby-252 datasets, on the profiles. The run\_id not only
defines the level of smoothing but also signifies the chosen height at
which the profiles were selected initially. Another important aspect is
the range of horizontal values (which is referred to as the y values in
\citet{aoas}) in the signatures. These have already been pre-processed
in the database to not include any grooves.

The idea therefore is to first use these raw values of the profile
directly in the chumbley algorithm, and see how the algorithm performs
for different coarseness values (smoothing parameter as referred in the
function lowess used in the chumbley algorithm).

\pagebreak

@@

We used the adjusted Chumbley method as proposed in \citet{hadler} and
implemented in the R package \texttt{toolmaRk} \citep{toolmark} on all
pairwise land-to-land comparisons of the Hamby scans (a total of 85,491
comparisons) with the pairwise sets for the comparisons given in the
table \ref{tab:param}.

\begin{table}[!h][!h]
\caption{\label{tab:param}Overview of parameter settings used for optimization and validation windows for bullet land signatures.}

\centering
\resizebox{\linewidth}{!}{\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrr}
\toprule
wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 120\\
wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 40 & 50 & 60\\
\bottomrule
\end{tabular}}}
\centering
\begin{tabular}[t]{lrrrrrrrrrrrrrrr}
\toprule
wo & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280\\
wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!h]
\caption{\label{tab:confusion}Confusion Table for different optimization window sizes with validation window size as 30.}

\centering
\begin{tabular}[t]{llrl}
\toprule
\hspace{1em}\hspace{1em}signif & match & Freq & Type\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Size of Optimization Window = 280}}\\
\hspace{1em}FALSE & FALSE & 78909 & True Negative\\
\hspace{1em}TRUE & FALSE & 4192 & False Positive (Type I)\\
\hspace{1em}FALSE & TRUE & 446 & False Negative (Type II)\\
\hspace{1em}TRUE & TRUE & 731 & True Positive\\
\bottomrule
\end{tabular}
\centering
\begin{tabular}[t]{llrl}
\toprule
signif & match & Freq & Type\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Size of Optimization Window = 120}}\\
\hspace{1em}FALSE & FALSE & 79249 & True Negative\\
\hspace{1em}TRUE & FALSE & 4523 & False Positive (Type I)\\
\hspace{1em}FALSE & TRUE & 386 & False Negative (Type II)\\
\hspace{1em}TRUE & TRUE & 854 & True Positive\\
\bottomrule
\end{tabular}
\centering
\begin{tabular}[t]{llrl}
\toprule
signif & match & Freq & Type\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Size of Optimization Window = 80}}\\
\hspace{1em}FALSE & FALSE & 79477 & True Negative\\
\hspace{1em}TRUE & FALSE & 4503 & False Positive (Type I)\\
\hspace{1em}FALSE & TRUE & 463 & False Negative (Type II)\\
\hspace{1em}TRUE & TRUE & 781 & True Positive\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Signatures}\label{signatures-1}

Figure \ref{fig:type2} gives an overview of type II error rates observed
when varying the window size in the optimization step. Two levels of
validation window size 30 and 50 were chosen as to compare the error
rates for different nominal type I errors. We notice that the trends for
these nominal type I errors are similar and in most cases a validation
window of 50 has higher type II error than for 30. A change in this
trend is seen for a 0.05 \(\alpha\) level, although the difference
between the two windows is very small for this case. We can also notice
an obvious trend of increase in the Type II error as the window of
optimization increases and see a minimum around the optimization window
size of 120 pixels. Hence we are inclined to choose a smaller validation
window size and optimization window as 120.

Table \ref{tab:confusion} shows the confusion tables with the
classification of type I and type II errors and how the numbers change
with a change in the optimization window. The windows represent areas to
the left of the window with minimum type II, near the minimum type II
window and to the far right of the minimum type II error

@@

Figure \ref{fig:type1} compares nominal (fixed) type I error and
actually observed type I errors for the parameter settings in table
\ref{tab:param}. With an increasing size of the window used in the
optimization step the observed type I error rate decreases (slighty).
This means as the optimization window increase the observed type I error
rate gets smaller. A smaller validation window on the other hand, tends
to be associated with a higher type I error rate. This can be better
imagined for a given window of optimization, where the actual Type I
error is comparable to the nominal level for only a select few
validation window sizes. For these comparable validation window sizes of
30 and 50 as done here, the actual type I error increases very slightly
and can be seen in Figure \ref{fig:type1}. This increase is not as much
when compared to the variation seen with the optimization window sizes.
This effect might be related to the increasing number of tests that fail
for larger optimization window sizes, in particular for non-matching
striae (see fig \ref{fig:missings}).

@@

\caption{Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$. The horizontal line in each facet indicates the nominal type I error rate.}\label{fig:type1}
\end{figure}

{\centering \includegraphics[width=\textwidth]{figures/wo120-1}

\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120}
\end{figure}

The actual type I error and type II error for signatures were also
compared for different validation window sizes. Figure \ref{fig:wo120}
shows the actual rates for different nominal \(\alpha\) levels. We can
see that the type II error rises with higher validation windows for the
smaller nominal \(\alpha\) levels while for the nominal \(\alpha\) =
0.05 its almost constant.

@@

\caption{Number of failed tests by the window optimization size, wo, and ground truth.}\label{fig:missings}
\end{figure}

@@

Figure \ref{fig:missings} gives an overview of the number of failed
tests, i.e.~tests in which a particular parameter setting did not return
a valid result. This happens, when the shift to align two signatures is
so large, that the remaining overlap is too small to accommodate windows
for validation. The problem is therefore exacerbated by a larger
validation window. Figure \ref{fig:missings} also shows that the number
of failed tests is approximately linear in the size of the optimization
window. Test results from different sources have a much higher chance to
fail, raising the question, whether failed tests should be treated as
rejections of the null hypothesis of same source. For known non-matches
there is a higher possibility that in the optimization pair of windows
where cross-correlations are maximum are too far apart, and same shifts
of this order hit the end of the signature.

\subsection{Profiles}\label{profiles-1}

Figure \ref{fig:prof_missings} (a) shows the type II error rates for
profiles for the optimization window 120 and validation window 30 with
varying level of coarseness. We can see that the type II error for all
the nominal \(\alpha\) levels is lowest in the range of 0.20 to 0.35.
Therefore, a value of 0.25 can be used keeping in mind it keeps the type
II error lowest while runnning simulations. Thus for comparisons of
different window sizes etc as seen in the differnt parts of Figure
\ref{fig:prof_missings} this coarseness value is used.

On the other hand Figure \ref{fig:prof_missings} (b) shows if the
coarseness level set in the chumbley agorithm has any effect on the
signatures, which are pre-processed and already smoothed to a certain
extent. From Figure \ref{fig:prof_missings} (b) we can notice that for
different nominal \(\alpha\) levels, the type II error fluctuates
slightly but does not change much, thereby helping us conclude that the
coarseness levels set in the LOWESS smoothing in the chumbley alggorithm
does effect the type II error much for signatures.

\subsubsection{Comparison of profiles and
signatures}\label{comparison-of-profiles-and-signatures}

Another reason for failed tests can be incorrect identification of
maximum correlation windows in the optimization step as seen in figure
\ref{fig:prof_missings}(d) because of the level of smoothing, as too
much smoothing would subdue intricate features that might otherwise help
in the correlation calculations and correct identification of maximum
correlation windows irrespective of the size. This would again cause a
simiar effect as explained for figure \ref{fig:missings} with validation
windows, irrespective of size, during the shifts end up at the ends of
the markings resulting in an invalid calculation and failed comparison
attempt.

In figure \ref{fig:prof_missings}(d) and (f), we compare profiles and
signatures on the basis of number of failed tests. The profiles chosen
for figure \ref{fig:prof_missings}(f) have a constant coarseness of 0.25
and window of optimization as 120. The signatures in this case are not
smoothed using the chumbley algorithm step of LOWESS smoothing. Instead
signatures are used as calculated by \citet{aoas}. The smoothing in
these signatures were determined and fixed on the basis of their
performance in the random forest based algorithm proposed by
\citet{aoas}. The comparison of profiles and signatures with variation
of validation window size therefore is made on even footing. The trends
are similar to figure \ref{fig:missings} in the sense that for known
non-matches the number of failed tests are more for both signatures and
profiles and increasing linearly with the validation window size. The
problem is however, worse for profiles which has higher number of failed
tests than signatures for all validation windows.

The total error for different validation window sizes for signatures and
profiles can be seen in figure \ref{fig:prof_missings} (e).The
optimization window size is 120 and profiles are calculated at a default
0.25 coarseness level while signatures as before are not smoothed again
in the modified chumbley algorithm. We can see that the total error is
always higher for profiles as compared to signatures for all sizes of
validation window.

{\centering \includegraphics[width=\textwidth]{figures/prof_missings-1}

\caption{Row 3:  Total error and Number of failed tests by the window validation size, wv, and ground truth, Row 2: Total error and Number of failed tests with Coarseness for both profiles and signatures, Row 1: Type II error for different coarseness levels as used in the modified chumbley algorithm for profiles and signatures}\label{fig:prof_missings}
\end{figure}

\pagebreak

\subsection{Conclusion}\label{conclusion}

The results suggest that the Nominal type I error \(\alpha\) value shows
dependence on the size of the window of optimization. For a given window
of optimization the actual Type I error is comparable to the nominal
level for only a select few validation window sizes and for comparable
validation window sizes of 30 and 50 as done here, the actual type I
error does not seem to vary as much as it varies with the optimization
window sizes . A Test Fail, i.e.~tests in which a particular parameter
setting did not return a valid result, happens, when the shift to align
two signatures is so large, that the remaining overlap is too small to
accommodate windows for validation, depends on whether known-match or
known non-matches has predictive value, with test results from different
sources having a much higher chance to fail. On conducting an analysis
of all known bullet lands using the adjusted chumbley algorithm, Type II
error was identified to be least bad for window of validation 30 and
window of optimization 120. In case of unsmoothed raw marks (profiles),
Type II error increases with the amount of smoothing and least for
LOWESS smoothing coarseness value about 0.25 or 0.3. In an effort to
identify the level of adaptiveness of the algorithm, comparisons were
made between signatures and profiles. Their comparison with respect to
validation window size for a fixed optimization window size suggested
that, profiles have a total error (i.e all incorrect classification of
known-matches and known non-matches) greater than or equal to the total
error of signatures for all sizes of validation window. Profiles also
fail more number of times than signatures in a test fail (for different
coarseness keeping windows fixed and also for different validation
windows keeping coarseness fixed) which lets us conclude that the
behaviour of the algorithm for the profiles instead of pre-processed
signatures is not better. Finally it should be noted that the current
version of the adjusted chumbley algorithm seems to fall short when
compared to other machine-learning based methods \citet{aoas}, and some
level of modification to the deterministic algorithm needs to be
identified and tested that would reduce the number of incorrect
classifications.

{\centering \includegraphics[width=\textwidth]{figures/wo120-profile-1}

\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120-profile}
\end{figure}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/failtest-coarseness-1}

\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-coarseness}
\end{figure}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/failtest-wv-1}

\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-wv}
\end{figure}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/failtest-wo-1}

\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-wo}
\end{figure}
